---
title: "lab2"
author: "Addison Schmidt"
date: "9/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The cocoa percents have a median of 70% and a similar mean of 71%. The distribution is not significantly skewed, and has a narrow 1st-3rd quartile; 
50% of the percentages fall between 70 and 74 percent cocoa. There are two outliers in the data, with cocoa percentages near 100%.
```{r}
library(classdata)
hist(choco$CocoaPercent)
boxplot(choco$CocoaPercent)
summary(choco$CocoaPercent)
```
2.  Burma is tied for the fewest with just 1 rating, while Venezuala has the most with over 200. Unsurprisingly, countries with more cocoa exports seem to tend towards
having more ratings. The first graph shows the 30 countries with the fewest, and the second the 31 with the most in the dataset.
```{r}
x = table(choco$CountryBeanOrigin)
x = sort(x)
barplot(x[1:30], las=2)
barplot(x[31:61], las=2)
```

3. The data show that there is a negative correlation between the cocoa percent and its rating. We are 95% certain that the correlation is between -0.18656 and -0.10912.
This is shown by the linear regression line on the scatterplot. The points themselves are difficult to read because they are numberous. When excluding chocolates with more
than 95% cocoa, we still find that this correlation holds - although to a lesser degree - yielding a 95% confidence interval from -0.10497 to -0.02577.
```{r}
cor.test(choco$CocoaPercent, choco$Rating)
plot(choco$CocoaPercent, choco$Rating)
model = lm(choco$Rating ~ choco$CocoaPercent, data=choco)
abline(model)

smaller = choco[choco$CocoaPercent < 95,]
cor.test(smaller$CocoaPercent, smaller$Rating)
plot(smaller$CocoaPercent, smaller$Rating)
model = lm(smaller$Rating ~ smaller$CocoaPercent, data=smaller)
abline(model)
```

4. Although there is variance in the mean, medians, modes, etc. of flavor ratings across countries as seen in the first graph, we do not see a significant difference in ratings
amongst countries with at least 100 ratings. The variace amongst other countries is likely driven by random variance, and is not statistically significant. Therefore, this
dataset does not show us a significant difference between countries.
```{r}
boxplot(Rating ~ CountryBeanOrigin ,data=choco)
most = choco[(x[choco$CountryBeanOrigin] > 100),]
boxplot(Rating ~ CountryBeanOrigin, data=most, las=2)
labels = c("Blend", "Dominican Republic", "Ecuador", "Madagascar", "Peru", "Venezuela")
for(label in labels) {
  temp = most[most$CountryBeanOrigin == label,]
  print(label)
  print(summary(temp$Rating))
}
```

5. I will first compare the company to CocoaPercent amongst the companies with the most reviews. The interesting part of this analysis is the variance within a company, as well as the differences between the cocoa amounts they use.
```{r}
tableCompany = table(choco$Company)
tableCompany = sort(tableCompany)
mostCompany = choco[(tableCompany[choco$Company] > 20),]
boxplot(CocoaPercent ~ Company, data=mostCompany, las=2)
labelsCompany = c("A. Morin", "Arete", "Bonnat", "Dandelion", "Domori", "Fresco", "Guitard", "Pralus", "Soma", "Valrhona", "Zotter")
for(label in labelsCompany) {
  tempCompany = mostCompany[mostCompany$Company == label,]
  print(label)
  print(summary(tempCompany$CocoaPercent))
}
```

From the summary statistics we see that the mean and median of all 11 of these companies is close to 70%. The boxplot shows us that most companies have a few outliers 
with higher cocoa percentage (likely specialized chocolates), and that most have a very small range from quartile 1 to quartile 3. Zotter is the exception to this, with a relatively much wider range of percentages.

Secondly, I will compare the review date by rating to see how the ratings have trended over time. The year 2021 will be omitted because reviews may vary by season, and 
not all seasons of 2021 have passed yet.
```{r}
choco = choco[choco$ReviewDate < 2021, ]
cor.test(choco$ReviewDate, choco$Rating)

x = c()
labels = c()
for(i in 2006:2020) {
  tmp = choco[choco$ReviewDate == i,]
  x[i-2005] = mean(tmp$Rating)
  labels[i-2005] = i
}
barplot(x, names.arg=labels, las=2)
boxplot(Rating ~ ReviewDate, data=choco, las=2)
```

While it's hard to see the trend in either the barplot or boxplot, we can construct a 95% confidence interval that the correlation is between 0.054 and 0.134. This is a small
positive correlation. From the boxplot, we can see that the interquartile range was greatest in 2006, and was lowest in 2017.